---
title: "Appendix"
output: pdf_document
---

```{r warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
library(Rtsne)
library(e1071)
library(glmnet)
load("data_nir_tablets.RData")
```
. 

Then we are checking for range of x before and after scaling to see how scaling impacts the data. 


```{r}
range(x)
dim(x)
x <- scale(x) #standardizing data
range(x)
rtsne <-Rtsne(x,perplexity =5)
cols <- c("black","darkorange2")[y+1]
plot(rtsne$Y, pch= 19, col = adjustcolor(cols,0.3))
```
I am initially scaling x (basically normalizing it for the given range). 


Data will be divided into 2 parts. The first part is the training part and the second part is the validation
set. I have divided them into the ratio of 80-20. We already have a test set which was provided. 
The primary purpose of a validation set is to tune in the value of the hyperparameter. 
```{r  warnings = FALSE}
TOT <- nrow(x)
N <- floor(TOT*0.8)
L <- floor(TOT*0.2)
B <- 100
table(y)
```
The below function is the misclass function which helps in matching the predicted output with the correct 
output. It helps in calculation of misclassification rate. 
```{r}

misclass_error <-function(y, yhat) 
  {
  tab <-table(y, yhat)
  1-classAgreement(tab)$diag
}
```


```{r}
tau <- 0.5
# tau is the threshold value for deciding between the outputs 0 and 1. 
# If the value is greater than tau, then 
# the output becomes 1, otherwise it stays 0. 
S <- 100 #length of the sequence of lambda that I will try. 
lambda <- seq(0.005,0.15,length = S)
# Getting a vector with all the lambda values that I have to try. 
# Creating the error train matrix of size B*S (number of trials times
# number of lambda value for each trial)
err_train <- err_val <- matrix(NA, B, S)
err_test <- lambda_best <- rep(NA, B)


#processing the data for each and every replication
for(b in 1:B){
  # sample train and validation data
train <- sample((1:TOT),N)
val <- setdiff(1:TOT,c(train))
  
  #train the model and save the model to fit variable
  fit <- glmnet(x[train,],y[train],family="binomial", alpha = 1,lambda = lambda)
  #using fit to train the model
  p_train <- predict(fit, newx = x[train,],type = "response")
  y_train <- apply(p_train, 2, function(v) ifelse(v > tau, 1, 0))
  #using the model for validation set
  p_val <-predict(fit,newx =x[val,],type ="response")
  y_val <-apply(p_val,2,function(v)ifelse(v>tau,1,0))
  #calculating training and validation error
  err_train[b,] <-sapply(1:S,function(s)misclass_error(y[train], y_train[,s])) 
  err_val[b,] <-sapply(1:S,function(s)misclass_error(y[val], y_val[,s]) )
  #putting the best result to lambda_best
  best <- which.min(err_val[b,])
  lambda_best[b] <- lambda[best]
}

```

```{r}
# take 1-lambda as lambda is inversely related to complexity
matplot(x =1-lambda,t(err_train),type ="l",lty =1,ylab ="Error",
        xlab ="1 -lambda",col =adjustcolor("black",0.1),log ="y")

# error on log scale
matplot(x =1-lambda,t(err_val),type ="l",lty =1,
        col =adjustcolor("deepskyblue2",0.1),add =TRUE,log ="y")
lines(1-lambda,colMeans(err_train),col ="black",lwd =2)
lines(1-lambda,colMeans(err_val),col ="deepskyblue3",lwd =2)
legend("topright",legend =c("Training error","Validation error"),
       fill =c("black","deepskyblue2"),bty ="n")
# get optimal lambda
lambda_star <-lambda[which.min(colMeans(err_val) ) ]
lambda_star
abline(v = 1 - lambda_star,col = "magenta")
```


```{r}
# compute misclassification error

fit <- glmnet(x[train,],y[train],family = "binomial",alpha=1,lambda = lambda_star)
# storing the final model that will be used for further calculation and on the test set. 
p_test <-predict(fit,newx =x[val,],type ="response")
xtest <-apply(p_test,2,function(v)ifelse(v>tau,1,0))
table(xtest,y[val])


misclass_error(xtest,y[val])

```

```{r}
x_test <- scale(x_test)
p_test <- predict(fit,newx = x_test,type = "response")
xtest <-apply(p_test,2,function(v)ifelse(v>tau,1,0))
table(xtest,y_test)
misclass_error(xtest,y_test)
```


